---
title: "Pentose_phosphate_pathway"
author: "CRISCI Emma"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction 

The systeme use is the systeme of the Pentose phosphate Pathway extract from the article Quantitative modeling of pentose phosphate
pathway response to oxidative stress reveals
a cooperative regulatory strategy from HURBAIN, Thammen and all

## Objectifs of the study

1. Modelisation of the Pentose Phosphate Pathway like in the reference article

#  Modelisation of the Pentose Phosphate Pathway like in the reference article

## Load of the external data about the reference system


```{r}
parameters <- read.table("parameter.txt", sep = " ", quote = "", header = FALSE, comment.char = ";")
colnames(parameters)<- c("parName","parValue","parUnit")
head(parameters)
```


```{r}
diffEquations <- read.table("differential_equations.txt", sep = " ", quote = "", header = FALSE,na.strings = "NA", fill = TRUE, comment.char = ";")
colnames(diffEquations)[1:2] <- c("diffEquaName", "equal")
head(diffEquations)
```

```{r}

```

