---
title: "try_equation"
author: "CRISCI Emma"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Fonctions utilisées
```{r}
# S. Charles (2022-03-10)
rm(list = ls())
library(tidyverse)
library(deSolve)
library(readxl)
library(ggplot2)
library(dplyr)
```


```{r}
                ## GENERAL FONCTION ##
# DIFFUSION LAW
# Function Diffusion to the compartment 1 to compartment 2
DIFF <- function(C1, C2, k, kdiff){
  # C1: matrix(time)
  # C2: matrix(time)
  # k: constant max
  # kdiff: constant diffusion
  phi <- k + kdiff * (C1-C2)
  return (phi)
}

# MASS ACTION LAW
# Function MA(m,n) (equation 5) Mass Action
MA <- function(sub, k, produit, inhib, Keq, Ki){
  # time: vector of time points
  # sub : matrix of substrat concentration (time)
  # produit : vector of n products concentration
  # inhib: vector of inhibitor
  # k: constant max
  # Km: vector of MM constants 
  # Keq: equilibrium constant
  # Ki: Inhibitor constant
  phi <- k * (prod(sub) - (prod(produit) / Keq)) / ( 1 + inhib/Ki )
  return(phi)
}
# /!\ value put by default for P, I, Keq, Ki, in case of the equation is not reversible and has no inhibitors

# MICHAELIS MENTEN LAW
MM <- function(sub, k, Km, produit, Keq, inhib, Ki){
  # time: vector of time points
  # sub : matrix of substrat concentration (time)
  # produit : vector of n products concentration
  # inhib: vector of inhibitor
  # k: constant max
  # Km: vector of MM constants
  # Keq: equilibrium constant
  # Ki: Inhibitor constant
  phi <- k * (prod(sub) - prod(produit)/Keq) / (1 + (prod(sub)+prod(produit))/Km + sum(inhib)/Ki)
  return(phi)
}
                ## PARTICULAR FONCTION ##
## GPx / GR

# MICHAELIS MENTEN LAW
MM_GPx_GR <- function(sub, k, Km, Un){
  # time: vector of time points
  # sub : matrix of substrat concentration (time)
  # k: constant max
  # Km: vector of MM constants
  # Un: if Un == 0 : GPx AND if Un == 1 : GR
  phi <- k * (prod(sub)) / (Un + sum(sub/Km)) # need to have the S in the same order than Km
  return(phi)
}

# MICHAELIS MENTEN LAW
MM_6PGD <- function(sub, k, Km, produit, inhib, Ki){
  # time: vector of time points
  # sub : matrix of substrat concentration (time)
  # produit : vector of n products concentration
  # inhib: vector of inhibitor
  # k: constant max
  # Km: vector of MM constants
  # Keq: equilibrium constant
  # Ki: Inhibitor constant
  phi <- k * (prod(sub))  / (1 + (prod(sub)+prod(produit))/Km + sum(inhib)/Ki)
  return(phi)
}

MA_GAPD <- function(sub, k, produit, inhib, Ki){
  # time: vector of time points
  # sub : matrix of substrat concentration (time)
  # produit : vector of n products concentration
  # inhib: vector of inhibitor
  # k: constant max
  # Km: vector of MM constants 
  # Keq: equilibrium constant
  # Ki: Inhibitor constant
  phi <- k * prod(sub) / ( 1 + inhib[1]/(Ki * (inhib[2]/ (inhib[2]+ 2*inhib[3]))))
  return(phi)
}

## Constant (need ?)
CONST <- function(k){
  return(k)
}
```

## Function study

### Diffusion
Variation of de [substrat]
```{r, eval=FALSE}
# Diffusion 
substrat <- seq(1,1000) # x
res <- sapply(substrat,DIFF,C2 = 5, k = 82.1486, kdiff=0.2016)#y

plot(x = substrat, y = res, main = "Exemple of the Diffusion function with Cext changing", xlab= "External Consentration", ylab= "Law rate dS/dt", las=1, type="l")
# Start condition
segments(x0 = 500, y0 = 0, x1 = 500, y1 = res[500], col = "red", lty = "dotted") 
segments(x0 = 0, y0 = res[500], x1 = 500, y1 = res[500], col = "red", lty = "dotted")
head(res)
rm(res, substrat)
```
Graph 1 : More the external concentration is more the coefficiant (Cext-Cin) is higher = quick the H202 go inside the cell. Like in this graph => always positif, we have always the H202 that go inside the cell.Red ligne show the initial condition chose by the article

Variation of le kdiff
```{r, eval=FALSE}
# Diffusion 
substrat <- seq(0.01, 0.4, by=0.01) # x
res <- sapply(substrat,DIFF,C1= 500, C2 = 5, k = 82.1486) # y

plot(x = substrat, y = res, main = "Exemple of the Diffusion function with k diff changing", xlab= "k diff", ylab= "Law rate dS/dt", las=1, type="l")
# Start condition
segments(x0 = 0.2016, y0 = 0, x1 = 0.2016, y1 = res[20], col = "red", lty = "dotted") 
segments(x0 = 0, y0 = res[20], x1 = 0.2016, y1 = res[20], col = "red", lty = "dotted")
head(res)
rm(res, substrat)
```
Graph 2 ; more k diff is higher, quick the H20 go inside the cell

### Mass action

Exemple with NNH who is MA(1,0)
```{r, eval=FALSE}
# MA
substrat <- seq(1,1000) # x
res <- sapply(substrat, MA, k = 4.2345 , produit = 0, inhib=0, Keq=1, Ki=1)#y

plot(x = substrat, y = res, main = "Exemple of the Diffusion function with concentration changing", xlab= "Consentration", ylab= "Law rate dS/dt", las=1, type="l")
head(res)
rm(res, substrat)
```

# Reactions
```{r, eval=FALSE}
# Example for phi_OX: T
# substrate: S = [H2O2]ext
k       <- 82.1486
kdiff   <- 0.2016
H202ext <- 500 # Only this metabolite consentration is a constant
S       <- 5 # arbitrary value chosen to try the MMs function
phi_OX <- DIFF(H202ext, S, k, kdiff)

# Example for phi_CAT: MA(1,0)
# substrate: S = [H2O2]
sub     <- 5 # arbitrary value chosen to try the MMs function 
produit <- 0 # No products
inhib   <- 0 # No inhibitors
k   <- 0.0879 # kCAT
Keq <- 1 # No
Ki  <- 1  # No inhibitors
phi_CAT <- MA(sub, k, produit, inhib, Keq, Ki)

# Example for phi_GPx: MM(2,0)
# substrates: S = [H2O2] and [GSH]  /  products: P = [GSSG] 
sub     <- c(2,5) # arbitrary value chosen to try the MMs function 
k <- 2.3736 # kGPx
Km <- 65    # KmGGR, KmNGR
sub <- 5 # arbitrary value chosen to try the MMs function
phi_PRPP <- MM_GPx_GR(sub, k, Km, Un=0)

print(phi_OX)
print(phi_CAT)
print(phi_PRPP)
```
# Modeling the system


## Differencial equation

matrice stochiometrie
```{r}
sys_sto_ode <- c(c(1,-1,-1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), c(0,0,-2,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), c(0,0,0,-1,1,-1,1,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0),c(0,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0),c(0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),c(0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), c(0,0,0,0,0,0,0,0,1,-1,-1,0,0,0,0,0,0,0,0,0,0,0,0), c(0,0,0,0,0,0,0,0,0,0,1,0,-1,0,-1,0,0,0,0,0,0,0,0), c(0,0,0,0,0,0,0,0,0,1,0,-1,-1,0,0,0,0,0,0,0,0,0,0), c(0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0,0,0), c(0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0,0), c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0),c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,1,-1,0,0,0,0,1),c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0,0),c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0),c(0,0,0,0,0,0,0,0,0,0,0,0,1,0,1,0,0,0,1,1,-1,0,-1),c(0,0,1,-1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),c(0,0,0,1,-1,1,-1,0,-1,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
name_equa_diff <- c("H2O2","GSH","NADPH","G6P","6PGL","6PG","Ru5P","X5P","R5P","S7P","E4P","GLC","F6P","FBP","DHAP","GAP","GSSG","NADPp")
name_phi       <- c("phi_OX","phi_CAT","phi_GPx","phi_GR","phi_NNH","phi_NHN","phi_G6PD","phi_6PGL","phi_6PGD","phi_RPI","phi_RPE","phi_PRPP","phi_TKT1","phi_TLD","phi_TKT2","phi_HK","phi_PGI","phi_PFK","phi_ALD","phi_TPI","phi_GAPD","phi_GLU","phi_TAL")

mat_stochio <- matrix(sys_sto_ode, nrow=23)

colnames(mat_stochio) <- name_equa_diff
rownames(mat_stochio) <- name_phi

mat_stochio_equadiff <- t(mat_stochio)

head(mat_stochio_equadiff)
```
import parameter table 
```{r}
parameters_table <- read_excel("parameters.xlsx")
parameters<-as.numeric(as.vector(t(parameters_table[,4])))
length(parameters)
```
Import initial consentration
```{r}
init_table <- read_excel("init.xlsx")
init<-as.numeric(as.vector(t(init_table[,2])))
length(init)
```


## Creation of the system
```{r}
syst <- function(t, y, parameters, msys){
  
  # PHI
  Phi <- matrix(c(
    DIFF(parameters[3], y[1], parameters[1], parameters[2]),
    MA(sub=y[1], k=parameters[4], produit=0, inhib=0, Keq=1, Ki=1),
    MM_GPx_GR(sub=c(y[1],y[2]), k=parameters[5], Km=c(parameters[6],parameters[7]), Un=0),
    MM_GPx_GR(sub=c(y[17],y[3]), k=parameters[8], Km=c(parameters[9],parameters[10]), Un=1),
    MA(sub=y[18], k=parameters[11], produit=0, inhib=0, Keq=1, Ki=1),
    MA(sub=y[18], k=parameters[12], produit=0, inhib=y[1], Keq=1, Ki=parameters[13]),
    MA(sub=c(y[4],y[18]), k=parameters[14], produit=0, inhib=y[3], Keq=1, Ki=parameters[15]),
    MA(sub=y[5], k=parameters[16], produit=0, inhib=0, Keq=1, Ki=1), 
    MM_6PGD(sub=c(y[6],y[18]), k=parameters[17], Km=parameters[18], produit=y[6], inhib=y[3], Ki=parameters[19]),
    MA(sub=y[7], k=parameters[20], produit=y[9], inhib=0, Keq=parameters[21], Ki=1), 
    MA(sub=y[7], k=parameters[22], produit=y[8], inhib=0, Keq=parameters[23], Ki=1),
    MM(sub=c(y[9]), k=parameters[24], Km=parameters[25], produit=0, Keq=1, inhib=0, Ki=1),
    MM(sub=c(y[9],y[8]), k=parameters[26], Km=parameters[27], produit=c(y[16],y[10]), Keq=parameters[28], inhib=0, Ki=1),
    MM(sub=c(y[16],y[10]), k=parameters[29], Km=parameters[30], produit=c(y[13],y[11]), Keq=parameters[31], inhib=0, Ki=1),
    MM(sub=c(y[11],y[8]), k=parameters[32], Km=parameters[33], produit=c(y[13],y[16]), Keq=parameters[34], inhib=0, Ki=1),
    MA(sub=y[12], k=parameters[35], produit=0, inhib=y[16], Keq=1, Ki=parameters[36]),
    MA(sub=y[4], k=parameters[37], produit=y[13], inhib=y[6], Keq=parameters[38], Ki=parameters[39]),
    MA(sub=y[13], k=parameters[40], produit=0, inhib=0, Keq=1, Ki=1) - MA(sub=y[14], k=parameters[41], produit=0, inhib=0, Keq=1, Ki=1),
    MA(sub=y[14], k=parameters[42], produit=c(y[15],y[16]), inhib=0, Keq=parameters[43], Ki=1),
    MA(sub=y[15], k=parameters[44], produit=y[16], inhib=0, Keq=parameters[45], Ki=1),
    MA_GAPD(sub=y[16], k=parameters[46], produit=0, inhib=c(y[1],y[2],y[17]), Ki=parameters[47]),
    parameters[49],
    MM(sub=c(y[10],y[16]), k=parameters[50], Km=parameters[52], produit=c(y[13],y[16]), Keq=parameters[51], inhib=0, Ki=1)
  ))
  dy <- msys %*% Phi
  
  
  return(list(as.vector(dy)))
}
```


# Modelisation
```{r}
time <- seq(0, 100, by=0.1)
solution <- lsoda(y = init, times = time, func = syst, parms = parameters, msys= mat_stochio_equadiff)
```


```{r}
length(solution)
```

log scale
```{r}
solution_log <- solution[,-c(1)]
```

model
```{r, echo=FALSE}
recupm <- read_excel("22-03-17_MODEL.xlsx")
```


```{r, echo=FALSE}
mmodel <- mutate_all(recupm, function(x) as.numeric(as.character(x))) # transform table into numeric values
```


```{r}
for (i in seq(2,length(name_equa_diff)+1)){
  plot(x=solution[,1], y=log10(solution[,i]), type = "l", main = paste("Concentration of", name_equa_diff[i-1]), xlab = "time (s)", ylab = "Concentration log",las=1)
  grid()
  }
```
```{r}
for (i in seq(2,length(name_equa_diff)+1)){
  plot(x=solution[,1], y=solution[,i],log="y", type = "l", main = paste("Concentration of", name_equa_diff[i-1]), xlab = "time (s)", ylab = "Concentration log",las=1)
  }
```



